\section{Additions}

The following constructs were added to those already defined in the \t{CAST.ml} given to us, of which all have been implemented and thoroughly tested: there are 127 test files, of which 126 have automatic testing. They amount to 2730+ lines of C\textminus\textminus, and result in 585+ test instances. They include complex tests: a number guessing game (\t{misc/guessnum.c}), a sudoku solver (\t{misc/sudoku\_solver.c}), a multithreaded word count (\t{misc/mwc.c}), and a sorting algorithm (\t{misc/rsort.c}).

\subsection{Unary operators}
\t{M\_DEREF}: accesses the address given by the expression.\\
Examples:
\begin{itemize}
    \item \t{*x} \to\ \ml{OP1(M_DEREF, VAR "x")}
    \item \t{*(x+1)} \to\ \ml{OP1(M_DEREF, OP2(S_ADD, VAR "x", CST 1))}
\end{itemize}

\t{M\_ADDR}: yields the address of an expression\\
Examples:
\begin{itemize}
    \item \t{\&x} \to\ \ml{OP1(M_ADDR, VAR "x")}
    \item \t{\&x[10]} \to\ \ml{OP1(M_ADDR, OP2(M_INDEX, VAR "x", CST 10))}
    \item \t{\&*x} \to\ \ml{OP1(M_ADDR, OP1(M_DEREF, VAR "x"))}
\end{itemize}
Errors:
\begin{itemize}
    \item \t{\&10}, \t{\&(x+1)}, \t{\&"abc"} \to\ `Indirection needs an lvalue'
\end{itemize}

Note:
\begin{itemize}
    \item \forall \t{e}, \ml{OP1(M_ADDR, OP1(M_DEREF, e))} is equivalent to \t{e}
    \item \forall \t{e} with an address, \ml{OP1(M_DEREF, OP1(M_ADDR, e))} is equivalent to \t{e}
\end{itemize}


\subsection{Binary operators}
The following binary operators were added:
\t{S\_SHL}, \t{S\_SHR}, \t{S\_OR}, \t{S\_XOR}, \t{S\_AND}
Examples:
\begin{itemize}
    \item \t{x << 2} \to\ \ml{OP2(S_SHL, VAR "x", CST 2)}
    \item \t{x >> 2} \to\ \ml{OP2(S_SHR, VAR "x", CST 2)}
    \item \t{x | 2} \to\ \ml{OP2(S_OR, VAR "x", CST 2)}
    \item \t{x \textasciicircum\ 2} \to\ \ml{OP2(S_XOR, VAR "x", CST 2)}
    \item \t{x \& 2} \to\ \ml{OP2(S_AND, VAR "x", CST 2)}
\end{itemize}

\subsection{Comparisons}
To simplify some expressions, the reduction step is allowed to perform the following transformations.\\
For any \t{a}, \t{b}:
\begin{itemize}
    \item \ml{EIF(CMP(C_EQ, a, b), 0, 1)} \to\ \ml{CMP(C_NE, a, b)}
    \item \ml{EIF(CMP(C_LE, a, b), 0, 1)} \to\ \ml{CMP(C_GT, a, b)}
    \item \ml{EIF(CMP(C_LT, a, b), 0, 1)} \to\ \ml{CMP(C_GE, a, b)}
\end{itemize}

\subsection{Extended assignment}
Examples:
\begin{itemize}
    \item \t{x += 2} \to\ \ml{OPSET(M_ADD, VAR "x", CST 2)}
    \item \t{x *= 2} \to\ \ml{OPSET(M_MUL, VAR "x", CST 2)}
    \item etc...
\end{itemize}
Errors:
\begin{itemize}
    \item \t{x []= 2} \to\ parsing error
    \item \t{2 += 2} \to\ `Extended assignment needs an lvalue'
\end{itemize}

\subsection{Control flow}
\t{CBREAK} (exit loop or switch), \t{CCONTINUE} (skip to next iteration of loop), \t{CTHROW} (raise exception)\\
Examples:
\begin{itemize}
    \item \t{break;} \to\ \ml{CBREAK}
    \item \t{continue;} \to\ \ml{CCONTINUE}
    \item \t{throw E(x);} \to\ \ml{CTHROW("E", VAR "x")}
    \item \t{throw E;} \to\ \ml{CTHROW("E", VAR "NULL")}
\end{itemize}
Errors:
\begin{itemize}
    \item \t{try \{ break; \}} \to\ `break may not reach outside of try'
    \item \t{try \{ continue; \}} \to\ `continue may not reach outside of try'
\end{itemize}

\subsection{Declarations}
\t{CLOCAL} is used to declare a local variable in the middle of a block.
Examples:
\begin{itemize}
    \item \t{int x;} \to\ \ml{CLOCAL[("x", None)]}
    \item \t{int x, y;} \to\ \ml{CLOCAL[("x", None), ("y", None)]}
    \item \t{int x = 1;} \to\ \ml{CLOCAL[("x", Some (CST 1))]}
\end{itemize}

\subsection{Switch}
\t{CSWITCH} declares a switch block.\\
Example:
\begin{minted}{c}
switch (x) {
    case 1: 1;
    case 2: 2;
    default: 3;
}
\end{minted}
parses to
\begin{minted}{ocaml}
CSWITCH (VAR x, [
    (1, CBLOCK [CEXPR (CST 1)]);
    (2, CBLOCK [CEXPR (CST 2)])
], CBLOCK [CEXPR (CST 3)])
\end{minted}

\subsection{Try}
\t{CTRY} declares an exception handling block.\\
Example:
\begin{minted}{c}
try { 1; }
catch (E x) { 2; }
catch (F _) { 3; }
catch (G) { 4; }
finally { 5; }
\end{minted}
parses to
\begin{minted}{ocaml}
CTRY (CBLOCK [CEXPR (CST 1)], [
    ("E", "x", CBLOCK [CEXPR (CST 2)]);
    ("F", "_", CBLOCK [CEXPR (CST 3)]);
    ("G", "_", CBLOCK [CEXPR (CST 4)])
], CBLOCK [CEXPR (CST 5)])
\end{minted}

\subsection{Other}
Some other features:
\begin{itemize}
    \item Variadic arguments are supported. They are declared with \t{...} then can be used with \t{va\_start} and \t{va\_arg}.
    \item Function pointers are supported. The original goal was to enable calling \t{qsort}.
    \item \t{switch} uses a jump table in the form of a binary tree. Its number of comparisons is logarithmic with regards to the number of \t{case} statements.
    \item \t{void} and \t{long} have been added as valid type declarations. They are equivalent to \t{int} but allow being more explicit on the purpose of a variable or function.
    \item The \t{assert} builtin has been implemented. Upon failure it raises an \t{AssertionFailure} exception which the user can catch.
    \item Color output is available by default, \t{--no-color} disables it.
    \item Expressions are reduced when possible by default, \t{--no-reduce} disables it.
\end{itemize}
